---
title: "Data analysis"
---
Here is a summary of all the variables in the dataset.
```{r, echo = F, warning = F, message = F}
library(tidyverse)
new_cars <- read.csv("used_cars_data.csv")
cars1 <- new_cars %>%
  separate(Mileage, into = c("Mileage_v", "Mileage_unit"), sep = " ")

cars2 <- cars1 %>%
  separate(Name, into = c("Brand", "Model"), sep = " ", extra = "merge") %>%
  mutate(Brand = case_when(
    Brand == "ISUZU" ~ "Isuzu",
    Brand == "Land" ~ "Land Rover",
    TRUE ~ Brand
  ))

cars <- cars2 %>%
  mutate("Engine/CC" = as.numeric(str_remove(Engine, "CC")),
         "Power/bhp" = as.numeric(str_remove(Power, "bhp")),
         "Mileage_value" = as.numeric(Mileage_v), 
         "Car_age" =  year(Sys.Date()) - Year
  ) %>%
  select(-c(Power, Engine, Mileage_v)) 

cars <- cars %>%
  select(Brand, Model, Location, Year, Car_age, Kilometers_Driven, Fuel_Type, Transmission, Owner_Type, Mileage_value, Mileage_unit, `Engine/CC`, `Power/bhp`, Seats, New_Price, Price)
knitr::kable(head(summary(cars)))
```
<br>

# 1. Lemon market: How high are the prices of second-hand cars in India?

```{r, echo = F, warning = F, message = F}
library(lubridate)
library(stringr)
cars %>%
  filter(Owner_Type == "Second") %>%
  mutate(Year = ym(str_c(Year, "-01"))) %>%
  na.omit() %>%
  group_by(Location, Year) %>%
  summarise(Price = mean(Price, na.rm = T)) %>%
  ggplot(aes(x = Year, y = Price)) +
  geom_line() +
  labs(title = "Prices of second hand cars in different parts of India against time") +
  facet_wrap(~ Location, nrow = 4, scales = "free_y") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 60, size = 7, vjust = 0.5))
```


```{r, echo = F, warning = F, message = F}
cars %>%
  filter(Owner_Type == c("First", "Second")) %>%
  mutate(Year = ym(str_c(Year, "-01"))) %>%
  na.omit() %>%
  group_by(Year, Owner_Type) %>%
  summarise(Price = mean(Price, na.rm = T)) %>%
  ggplot(aes(x = Year, y = Price)) +
  geom_line() +
  geom_point(colour = "red") +
  facet_wrap(~ Owner_Type) +
  labs(title = "Overall prices of first and second hand cars in India against time") +
  theme_bw() 
```
<br>

There has been an increase in second-hand car prices till 2017 (though this may vary, eg. Bangladore and Delhi experienced increases earlier) but seems to be mostly coming down after that. Perhaps the problem of the lemon market is slowly being alleviated as the difference in first and second hand car prices are increasing.

# 2. Carbon emissions? Average kilometers driven in India

One outlier from Chennai removed
```{r, echo = F, warning = F, message = F}
cars %>%
  filter(Kilometers_Driven <= 4e+06) %>%
  mutate(Year = ym(str_c(Year, "-01"))) %>%
  na.omit() %>%
  group_by(Year, Location) %>%
  summarise(Ave_km_driven = mean(Kilometers_Driven, na.rm = T)) %>%
  ggplot(aes(x = Year, y = Ave_km_driven)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  labs(title = "Graph of average kilometers driven in each part of India against time",
       x = "Year",
       y = "Average kilometers driven")
```
<br>
```{=html}
<iframe height="400" width="100%" frameborder="no" src="https://cc01lz-yasmineong.shinyapps.io/MyShinyApp/"> </iframe>
```

# 3. Preferences and car usage in India 

```{r, echo = F, warning = F}
library(patchwork)
plot1 <- cars %>%
  ggplot(aes(x = Fuel_Type)) +
  geom_bar(fill = "#0AE1FB") +
  labs(title = "Count of cars per fuel type",
       x = "Fuel type",
       y = "Count") +
   theme(plot.title = element_text(size = 10),
         axis.title.x = element_text(size = 8),
         axis.title.y = element_text(size = 8))

plot2 <- cars %>%
  ggplot(aes(x = Transmission)) +
  geom_bar(fill = "red") +
  labs(title = "Count of cars per transmission type",
       x = "Transmission type",
       y = "Count") +
   theme(plot.title = element_text(size = 10),
         axis.title.x = element_text(size = 8),
         axis.title.y = element_text(size = 8))

plot3 <- cars %>%
  ggplot(aes(x = Owner_Type)) +
  geom_bar(fill = "#0AFBBD") +
  labs(title = "Count of cars per owner type",
       x = "Owner type",
       y = "Count") +
   theme(plot.title = element_text(size = 10),
         axis.title.x = element_text(size = 8),
         axis.title.y = element_text(size = 8))

plot4 <- cars %>%
  ggplot(aes(x = Brand)) +
  geom_bar(fill = "#FAA4E8") +
  labs(title = "Count of cars per brand",
       x = "Brands",
       y = "Count") +
   theme(plot.title = element_text(size = 10),
         axis.title.x = element_text(size = 8),
         axis.title.y = element_text(size = 8),
         axis.text.x = element_text(angle = 45, size = 5, vjust = 0.5))

plot5 <- cars %>%
  count(Model) %>%
  arrange(desc(n)) %>%
  head(10) %>%
  ggplot(aes(x = Model, y = n)) +
  geom_col(fill = "Black") +
  labs(title = "Count of cars per (top 10) model",
       x = "Models",
       y = "Count") +
   theme(plot.title = element_text(size = 10),
         axis.title.x = element_text(size = 8),
         axis.title.y = element_text(size = 8),
         axis.text.x = element_text(angle = 45, size = 7, vjust = 0.5))

plot6 <- cars %>%
  ggplot(aes(x = Location)) +
  geom_bar(fill = "#F5B041") +
  labs(title = "Count of cars per location",
       x = "Location",
       y = "Count") +
   theme(plot.title = element_text(size = 10),
         axis.title.x = element_text(size = 8),
         axis.title.y = element_text(size = 8),
         axis.text.x = element_text(angle = 45, size = 7, vjust = 0.5))

plot1 + plot2 + plot3 + plot4 + plot5 + plot6 + plot_layout(ncol = 2) 
```
<br>

```{r, echo = F}
library(patchwork)
market_share <- cars %>%
  group_by(Brand) %>%
  summarise(count = n()) %>%
  mutate(market_share = count / sum(count) * 100) %>%
  arrange(desc(market_share)) %>%
  head(10)
pie(market_share$market_share, 
    labels = paste(market_share$Brand, ": ", round(market_share$market_share, 2), "%"), 
    main = "Market Share Analysis by Car Brand (Top 10)",
    col = c("#004BF8", "#0080F8", "#00A9F8", "#02CEF7", "#02E8F7", "#A3F7FF", "#A3AEFF", "#C489FF", "#B603FF", "#6C03BE"),
    font = 3,
    cex = 0.8)
```
```{r, echo = F, message = F, warning = F}
cars_summary <- cars %>%
  group_by(Brand) %>%
  summarise(Count = n()) %>%
  arrange(desc(Count))

top_10_brands <- cars_summary %>%
  head(10) %>%
  pull(Brand)

cars$Top_10 <- ifelse(cars$Brand %in% top_10_brands, "Top 10", "Others")

ggplot(cars, aes(x = Price, y = Brand)) +
  geom_boxplot() +
  facet_wrap(~ Top_10, scales = "free_y") +
  labs(title = "Price Ranges for Each Car Brand", 
       x = "Price", 
       y = "Brand") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
```
<br>

Maruti seems to be the most popular car brand that Indian consumers go for. Upon further inspection into the prices of the individual car brands, we seem to see a preference for the lowest costing car (Maruti) among the top 10.

```{r, echo = F, warning = F, message = F}
library(patchwork)
graph1 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = Price, fill = Transmission)) +
  geom_density(adjust = 0.5) 

graph2 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = Car_age, fill = Transmission)) +
  geom_density(adjust = 0.5) 

graph3 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = Kilometers_Driven, fill = Transmission)) +
  geom_density(adjust = 0.5) +
   theme(axis.text.x = element_text(size = 6, angle = 25))

graph4 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = Mileage_value, fill = Transmission)) +
  geom_density(adjust = 0.5) 

graph5 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = `Engine/CC`, fill = Transmission)) +
  geom_density(adjust = 0.5) 

graph6 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = `Power/bhp`, fill = Transmission)) +
  geom_density(adjust = 0.5) +
  theme(axis.text.x = element_text(size = 6))

graph7 <- cars %>%
  na.omit() %>%
  ggplot(aes(x = Seats, fill = Transmission)) +
  geom_density(adjust = 0.5) 

graph1 + graph2 + graph3 + graph4 + graph5 + graph6 + graph7 + plot_layout(ncol = 2)
```

1. Preferences for low costs, 5-seater vehicles, lower power and engine vehicles.
2. Automatic cars seem to have higher power and engine than manual cars --> More may opt for manual cars since they are of lower power and engine.
3. Cars in India do not travel very long distances.
4. Most of the histograms are highly skewed, which may lead to some poor performances of certain models.

# 4. Predictors in the market

```{r, echo = F}
library(ggcorrplot)
numerical_variables <- cars %>%
  na.omit() %>%
  select(Price, Car_age, Kilometers_Driven, Mileage_value, `Engine/CC`, `Power/bhp`, Seats) 

ggcorrplot(cor(numerical_variables),
           hc.order = TRUE,  
           lab = TRUE,
           lab_size = 2,
           outline.color = "black",  
           colors = c("blue", "white", "red"),  
           title = "Correlation Heatmap of Numerical variables")

```

1. Strongest correlation between the power and engine capacity of a car.

2. There is a strong correlation between costs and power of the car --> the higher the BHP, the higher the costs of the car which matches what vehicle analytics say. However, since customers in India tend to go for lower prices, producers need not create so many high-powered cars. Moreover, India's cities are known for their traffic congestion thus smaller, less powerful engines are practical for navigating these conditions efficient.

3. Relationship of price and engine of a car is also moderately strong. This seems to be in line with India's taxation policies: they impose higher taxes on cars with larger engines and as a result higher BHP, making such vehicles much more unafforable.

Link for reference: https://www.cars24.com/blog/what-is-bhp-in-cars/